---
title: "Covariate rasterization"
author: "Mikaella Kontopoulou"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    number_sections: true
    code_folding: show 
    css: "styles.css"
---

# Load packages
```{r load-packages}
suppressMessages(suppressWarnings(library(haven))) 
suppressMessages(suppressWarnings(library(broom)))
suppressMessages(suppressWarnings(library(data.table)))
suppressMessages(suppressWarnings(library(DescTools)))
suppressMessages(suppressWarnings(library(fields)))
suppressMessages(suppressWarnings(library(foreign)))
suppressMessages(suppressWarnings(library(MASS)))
suppressMessages(suppressWarnings(library(naniar)))
suppressMessages(suppressWarnings(library(plyr)))
suppressMessages(suppressWarnings(library(raster)))
suppressMessages(suppressWarnings(library(readstata13)))
suppressMessages(suppressWarnings(library(rms)))
suppressMessages(suppressWarnings(library(spdep)))
suppressMessages(suppressWarnings(library(summarytools)))
suppressMessages(suppressWarnings(library(tidyverse)))
suppressMessages(suppressWarnings(library(lubridate)))
suppressMessages(suppressWarnings(library(zoo)))
suppressMessages(suppressWarnings(library(stringi)))
suppressMessages(suppressWarnings(library(leaflet)))
suppressMessages(suppressWarnings(library(stars)))
suppressMessages(suppressWarnings(library(htmlwidgets)))
suppressMessages(suppressWarnings(library(sf)))
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(purrr)))
suppressMessages(suppressWarnings(library(readxl)))
suppressMessages(suppressWarnings(library(raster)))
suppressMessages(suppressWarnings(library(countrycode)))
suppressMessages(suppressWarnings(library(scales)))
suppressMessages(suppressWarnings(library(terra)))
suppressMessages(suppressWarnings(library(fs)))
suppressMessages(suppressWarnings(library(bit64)))
suppressMessages(suppressWarnings(library(stringr)))
suppressMessages(suppressWarnings(library(RMariaDB)))
```

# Import DHS IR & MR data
DHS IR/MR files are restricted-access and cannot be redistributed. Approved users must download them directly from The DHS Program and run this code locally.

# Variable selection pre-requisites
```{r pre-select}
Dataframes <- ls()[sapply(ls(), function(x) inherits(get(x), c("data.frame", "tbl_df")))] 
Geolocations <-  Dataframes[grepl("^[A-Z]{2,3}[G]$", Dataframes)]
Dataframes <- Dataframes[grepl("^[A-Z]{2,3}[FM]$", Dataframes)]

Variables <- c("V001", "V130", "V131", "MV001", "MV130", "MV131")

Names <- c(V001 = "ID",
           V130 = "Religion",
           V131 = "Ethnicity",
           MV001 = "ID",
           MV130 = "Religion",
           MV131 = "Ethnicity")

MissingVariables <- function(df, Column) {MissColumn <- setdiff(Column, names(df))
                                          df[MissColumn] <- NA
                                          return(df)}

ReplaceValue4Label <- function(df) {for(Column in names(df)) {if(!is.null(attr(df[[Column]], 
                                                                               "labels"))) {labels <- attr(df[[Column]], "labels")
                                                                                            df[[Column]] <- names(labels)[match(df[[Column]], labels)]}}
                                                              return(df)}
```

# Variable selection
```{r select}
for (df in Dataframes) {DF <- get(df) %>%
                              dplyr::select(all_of(intersect(Variables, names(.)))) %>%
                              rename_with(~ Names[.x]) %>%
                              MissingVariables(unique(Names)) %>%
                              ReplaceValue4Label() %>%
                              mutate(DF = paste0(substr(df, 1, nchar(df) - 1)))
                        
                        assign(df, DF)}

for (df in Geolocations) {DF <- get(df) %>%
                                dplyr::select("DHSCLUST", "LATNUM", "LONGNUM", "geometry") %>%
                                dplyr::rename(ID = DHSCLUST,
                                              x = LONGNUM,
                                              y = LATNUM) %>%
                                st_drop_geometry() %>%
                                mutate(DF = paste0(substr(df, 1, nchar(df) - 1)))
                        
                        assign(df, DF)}

rm(Dataframes, Geolocations, Names, Variables, MissingVariables, ReplaceValue4Label, df, DF)
```

# Data merge
```{r merge}
Data <- bind_rows(lapply(setdiff(ls()[!grepl("^[A-Za-z]{2}G$", ls())], 
                                 c("Covariates", "params")), 
                         function(x) {df <- get(x)
                                      df[] <- lapply(df, as.character)
                                      return(df)}))

Geodata <- bind_rows(lapply(ls()[grepl("^[A-Za-z]{2}G$", ls())], function(x) get(x)))

rm(list = setdiff(ls(), c("Data", "Geodata")))
```

# Recoding
```{r recode}
Data <- Data %>%
        mutate(Judaism = ifelse(Religion %in% c("Jewish", "Zionist", "Zion"), 1,
                         ifelse(is.na(Religion), NA, 0)),
               Hinduism = ifelse(Religion %in% c("Hindu", "Hinduism"), 1, 
                          ifelse(is.na(Religion), NA, 0)),
               Islam = ifelse(Religion %in% c("Islam", "Islamic", "Moslem", "Muslim", "Muslin", "Bektashi", "Musulman"), 1,
                       ifelse(is.na(Religion), NA, 0)),
               Buddhism = ifelse(Religion %in% c("Buddhism", "Buddhist", "Buddhist / Neo-Buddhist", "Budhist"), 1, 
                          ifelse(is.na(Religion), NA, 0)),
               Atheism = ifelse(Religion %in% c("Atheist", "Aucune", "no religion", "No religion", "No religion/atheists",
                                                "No religion/none", "None", "Not religious", "Refused/DK", "DK", "Sans religion",
                                                "Without religion"), 1,
                         ifelse(is.na(Religion), NA, 0)),
               Christianity = ifelse(Religion %in% c("Adventist", "Adventiste", "Aglipay", "Anglican", "Anglican Church", 
                                                     "Armée de salut", "Methodist", "Assembly of god", "Assembly of God", 
                                                     "Apostolic sect", "Autres chretiens", "Autres Protestants", "Baptist", 
                                                     "Charismatic", "Christian", "Christian/protesstan", "Kimbanguiste",
                                                     "Christian/protestant", "Christianity", "Elcin", "Evangelic", 
                                                     "Evangelical", "Evangelical/pentecostal", "Evangelical / Protestant",
                                                     "Evangelical churches", "Evangelical presbyterian", "FJKM/FLM/Anglikana", 
                                                     "Iglesia ni Cristo", "Jehovah's Witness", "Jehovah's witnesses", 
                                                     "Jehovah witness", "Jehovah Witness", "Lesotho Evangelical church", 
                                                     "Other christian", "Other Christian", "Other christians", "Other chritians",
                                                     "Other Christian (not otherwise categorisable)", "Other christian (specify)", 
                                                     "Other Christian religion", "Other Christians", "Other Protestants", "CCAP", 
                                                     "Pentecostal", "Pentecostal/Born Again/Evangelical", "Pentecotist", "Celestes",
                                                     "Pentecostal/charismatic", "Presbyterian", "Prostestan", "Protestant", 
                                                     "Protestant Methodist", "Protestant/Anglican", "Seventh-day adventist", 
                                                     "Roman Catholic", "Salvation Army", "Roman Catholic church", "Roman catholic",
                                                     "Seventh Day Adventist", "Seventh Day Adventist / Baptist", "Revival church",
                                                     "Protestant/Methodist/Adventist/Jehova Witness", "Protestante Methodiste", 
                                                     "Protestant (ex. evangelical, baptist, jehovah witness)", "African instituted churches",
                                                     "Cathlic", "Catholic", "Catholic/Greek Cath", "Catholic/greek cath.", "Catholique",
                                                     "Chretien", "Evangelical Presbyterian", "Jehovah’s witness", "Lesotho Evangelical Church",
                                                     "Non denominational/Charismatic/Apostolic", "Orthodox", "Pentecostal/born again/evangelical",
                                                     "Salvation army", "Seven days/Baptist", "Seventh day adventist"), 1, 
                               ifelse(is.na(Religion), NA, 0)),
               FolkR = ifelse(Religion %in% c("Animalist", "Animist", "Other traditional", "Traditionists", "Traditional/animist", 
                                              "Traditional/Animist", "Tradition/animist", "Traditional", "Traditional religion",
                                              "Traditional/spiritualist", "Traditionalist", "Autre traditionnelles", 
                                              "Indigenous spirituality", "Kirat", "Bundu dia kongo", "Vodoun", "Jain", "Confucian",
                                              "Animist/Traditional religion", "Animiste", "Traditionalist / Animist", "animist"), 1, 
                       ifelse(is.na(Religion), NA, 0)),
               OtherR = ifelse(Religion %in% c("Mammon", "Baha'i", "Vaudousant", "Vuvamu", "Other", "Autres religions", "Khong Hu Chu", 
                                               "Other religions", "Parsi / Zoroastrian", "Sikh", "Sect", "Universal", "Others",
                                               "Rastafarian"), 1, 
                        ifelse(is.na(Religion), NA, 0)),
               Ethnicity = gsub(" and.*", "", gsub("\\s*/\\s*", "/", stri_trans_general(tolower(Ethnicity), "Latin-ASCII")))) %>%
        mutate(Ethnicity = ifelse(grepl("other", Ethnicity, ignore.case = TRUE), "other", 
                           ifelse(Ethnicity %in% c("don't know", "dk, unsure", "refused/not stated", "missing", "refused to say") , NA, 
                           ifelse(Ethnicity %in% c("dk/none", "no caste/tribe"),  "None", 
                           ifelse(Ethnicity %in% c("non-gambian", "non arab nationalities", "foreign/non-Congolese", "not-ivorian",
                                                   "not a Senegalese", "foreign language", "foreign/non-congolese", "not a senegalese"), "other", 
                           ifelse(Ethnicity %in% c("black/african", "black/mulato/afro-colombian/afro-descendent"), "african", 
                           ifelse(Ethnicity %in% c("indian/asian"),  "indian", 
                           ifelse(Ethnicity %in% c("roma (gypsy)", "gypsy (rom)"),  "gypsy", Ethnicity)))))))) %>%
        dplyr::select(DF, ID, Judaism, Hinduism, Islam, Buddhism, Atheism, Christianity, FolkR, OtherR, Ethnicity)
```

# Seperate INDIA data for independent processing
```{r india-subset}
IAD <- Data %>% filter(DF == "IA")
IAG <- Geodata %>% filter(DF == "IA")

Data <- Data %>% filter(DF != "IA")
Geodata <- Geodata %>% filter(DF != "IA")
```

# Compute Ethnic Fractionalization Index [EFI]
```{r efi}
CountPerE <- Data %>%
             group_by(DF, ID, Ethnicity) %>%
             summarise(Count = n(), .groups = 'drop') 

NPerG <- CountPerE %>%
         group_by(DF, ID) %>%
         summarise(N = sum(Count), .groups = 'drop') 

EFI <- merge(CountPerE, NPerG, by = c("DF", "ID")) %>%
       mutate(ProportionPerE = Count / N) %>%
       group_by(DF, ID) %>%
       summarise(SProportionPerE = sum(ProportionPerE^2), .groups = 'drop') %>%
       mutate(EFI_prop = round(1 - SProportionPerE, 4)) %>%
       dplyr::select(-SProportionPerE)

rm(CountPerE, NPerG)
```

# Compute Religion Diversity Index [RDI]
```{r rdi}
RDI <- Data %>%
       dplyr::select(-Ethnicity) %>%
       dplyr::group_by(DF, ID) %>%
       dplyr::summarise(N = n(),
                        Judaism_percent = round(sum(Judaism, na.rm = TRUE) / N, 4)*100,
                        Hinduism_percent = round(sum(Hinduism, na.rm = TRUE) / N, 4)*100,
                        Islam_percent = round(sum(Islam, na.rm = TRUE) / N, 4)*100,
                        Buddhism_percent = round(sum(Buddhism, na.rm = TRUE) / N, 4)*100,
                        Atheism_percent = round(sum(Atheism, na.rm = TRUE) / N, 4)*100,
                        Christianity_percent = round(sum(Christianity, na.rm = TRUE) / N, 4)*100,
                        OtherR_percent = round(sum(OtherR, na.rm = TRUE) / N, 4)*100,
                        FolkR_percent = round(sum(FolkR, na.rm = TRUE) / N, 4)*100,
                        .groups = 'drop') %>%
       mutate(SumPercentPerR = Judaism_percent^2 + Hinduism_percent^2 + 
                                Islam_percent^2 + Buddhism_percent^2 + 
                                Atheism_percent^2 + Christianity_percent^2 + 
                                OtherR_percent^2 + FolkR_percent^2,
              InvertedSumPercentPerR = 100^2 - SumPercentPerR,
              EqualPercentPerR = 100^2 - (SumPercentPerR/8),
              RDI_prop = ifelse(SumPercentPerR == 0, NA, InvertedSumPercentPerR / EqualPercentPerR)) %>%
      dplyr::select(DF, ID, RDI_prop)
```

RDI: (100^2 - SUM(Judaism_prop^2 + Hinduism_prop^2 + Islam_prop^2 + Buddhism_prop^2 + Atheism_prop^2 + Christianity_prop^2 + OtherR_prop^2)/
     (100^2 - (SUM(Judaism_prop^2 + Hinduism_prop^2 + Islam_prop^2 + Buddhism_prop^2 + Atheism_prop^2 + Christianity_prop^2 + OtherR_prop^2)/8))

# Merge data
```{r merge2}
Data <- merge(EFI, RDI, by = c("DF", "ID"))

rm(EFI, RDI)
```

# Add geolocation
```{r geolocate}
Data <- merge(Data, Geodata, by = c("DF", "ID"))

rm(Geodata)
```

# Split & Remove NAs
```{r subset}
Dataframes <- vector()

for (country in unique(Data$DF)) {DF <- Data[Data$DF == country, ]
                                  for (Column in grep("_prop$", names(DF), value = TRUE)) {
                                       if (all(is.na(DF[[Column]]))) {Name <- sub("_prop$", "", Column)
                                                                      Remove <- grep(paste0("^", Name), names(DF), value = TRUE)
                                                                      DF <- DF[ , !(names(DF) %in% Remove)]}}
                                  assign(paste0(country), DF)
                                  Dataframes <- c(Dataframes, paste0(country))}

rm(Column, country, Name, Remove)
```
        
# INDIA independent pre-processing
```{r india-process}
sf_use_s2(FALSE)

ADM2 <- st_read("Input/ADM2.gpkg") %>% 
        dplyr::select(geom) %>%
        rename("geometry" = geom) %>%
        mutate(centroid = st_centroid(geometry)) 

sf_use_s2(TRUE)

st_geometry(ADM2) <- "geometry"

IAG <- st_as_sf(IAG, coords = c("x", "y"), crs = 4326) 

sf_use_s2(FALSE)

IAG <- st_join(IAG, ADM2, join = st_within) %>%
       st_drop_geometry() 

sf_use_s2(TRUE)

st_geometry(IAG) <- "centroid"

IAG <- merge(IAD, IAG, by = c("DF", "ID")) %>% group_by(centroid) %>% mutate(ID = cur_group_id()) %>% ungroup() %>% st_as_sf()

IAG <- IAG[!st_is_empty(IAG$centroid),]

IAG$x <- st_coordinates(IAG)[,1]
IAG$y <- st_coordinates(IAG)[,2]

IAG <- IAG %>% st_drop_geometry()

rm(IAD, ADM2)
```

# Cluster aggregation for INDIA
```{r india-aggregate}
# Compute Ethnic Fractionalization Index [EFI]
CountPerE <- IAG %>%
             group_by(DF, ID, Ethnicity) %>%
             summarise(Count = n(), .groups = 'drop') 

NPerG <- CountPerE %>%
         group_by(DF, ID) %>%
         summarise(N = sum(Count), .groups = 'drop') 

EFI <- merge(CountPerE, NPerG, by = c("DF", "ID")) %>%
       mutate(ProportionPerE = Count / N) %>%
       group_by(DF, ID) %>%
       summarise(SProportionPerE = sum(ProportionPerE^2), .groups = 'drop') %>%
       mutate(EFI_prop = round(1 - SProportionPerE, 4)) %>%
       dplyr::select(-SProportionPerE)

# Compute Religion Diversity Index [RDI]
RDI <- IAG %>%
       dplyr::select(-Ethnicity) %>%
       dplyr::group_by(DF, ID) %>%
       dplyr::summarise(N = n(),
                        Judaism_percent = round(sum(Judaism, na.rm = TRUE) / N, 4)*100,
                        Hinduism_percent = round(sum(Hinduism, na.rm = TRUE) / N, 4)*100,
                        Islam_percent = round(sum(Islam, na.rm = TRUE) / N, 4)*100,
                        Buddhism_percent = round(sum(Buddhism, na.rm = TRUE) / N, 4)*100,
                        Atheism_percent = round(sum(Atheism, na.rm = TRUE) / N, 4)*100,
                        Christianity_percent = round(sum(Christianity, na.rm = TRUE) / N, 4)*100,
                        OtherR_percent = round(sum(OtherR, na.rm = TRUE) / N, 4)*100,
                        FolkR_percent = round(sum(FolkR, na.rm = TRUE) / N, 4)*100,
                        .groups = 'drop') %>%
       mutate(SumPercentPerR = Judaism_percent^2 + Hinduism_percent^2 + 
                                Islam_percent^2 + Buddhism_percent^2 + 
                                Atheism_percent^2 + Christianity_percent^2 + 
                                OtherR_percent^2 + FolkR_percent^2,
              InvertedSumPercentPerR = 100^2 - SumPercentPerR,
              EqualPercentPerR = 100^2 - (SumPercentPerR/8),
              RDI_prop = ifelse(SumPercentPerR == 0, NA, InvertedSumPercentPerR / EqualPercentPerR)) %>%
      dplyr::select(DF, ID, RDI_prop)

IA <- merge(EFI, RDI, by = c("ID", "DF"))
IA <- merge(IA, (IAG %>% dplyr::select(DF, ID, x, y)), by = c("DF", "ID")) %>%
      distinct()

rm(CountPerE, NPerG, EFI, RDI, IAG)
```

# Krigging pre-requisites
```{r pre-krig}
set.seed(1212)

Dataframes <- c(Dataframes, "IA")

Kriggdata <- list()

for (df in Dataframes) {DF <- get(df)

                        Trainset <- list()
                        Testset <- list()
                        TrainCC <- list()
                        TestCC <- list()
  
                        for (Variable in c("EFI", "RDI")) {
                             Column <- paste0(Variable, "_prop")
                             if (Column %in% names(DF)) {NotNA <- DF[!is.na(DF[[Column]]), ]
                                                         Index <- sample(seq_len(nrow(NotNA)), ceiling(0.2 * nrow(NotNA)), replace = FALSE)
                                                         TrainCC[[Variable]] <- NotNA[-Index, (ncol(NotNA)-1) : ncol(NotNA)]
                                                         TestCC[[Variable]] <- NotNA[Index, (ncol(NotNA)-1) : ncol(NotNA)]
                                                         Trainset[[Variable]] <- NotNA[[Column]][-Index] # [80% of data]
                                                         Testset[[Variable]] <- NotNA[[Column]][Index]}} # [20% of data]
                       
                        Kriggdata[[df]] <- list(TrainCC = TrainCC,
                                                TestCC = TestCC,
                                                Trainset = Trainset,
                                                Testset = Testset)}

rm(Column, Index, Variable, NotNA, df, DF, Testset, TrainCC, TestCC, Trainset)
```

# Krigging function construction for proportions [Must be 0-1 bound]  
```{r kreg}
CompareSP1 <- function(TRAINX, 
                       TESTX, 
                       TRAINCOORDS,
                       TESTCOORDS,
                       dd_values) {results <- data.frame(dd = dd_values, 
                                                         MAE = numeric(length(dd_values)), 
                                                         RMSE = numeric(length(dd_values)))
 
                                   for (i in 1:length(dd_values)) {dd <- dd_values[i]
                                                                   # Adjustment to extreme values to prevent errors
                                                                   if (any(TRAINX == 0) | any(TRAINX == 1)) {TRAINX[TRAINX == 0] <- 0.0001
                                                                                                             TRAINX[TRAINX == 1] <- 0.9999}
                                                                   # Logit transformation 
                                                                   logitTRANS <- log(TRAINX / (1 - TRAINX))
                                                                                                           
                                                                   # Predict
                                                                   dists <- as.numeric(summary(dist(TRAINCOORDS))[dd])
                                                                   krig.fit <- Krig(TRAINCOORDS, logitTRANS, theta = dists)
                                                                   pred.krig <- predict(krig.fit, TESTCOORDS)
                                                                   backtrans <- exp(pred.krig) / (1 + exp(pred.krig))
                                                                                                           
                                                                   # Compute MAE & RMSE for validation
                                                                   MAE <- mean(abs(backtrans - TESTX))
                                                                   RMSE <- sqrt(mean((backtrans - TESTX)^2))
                                                                   results$MAE[i] <- MAE
                                                                   results$RMSE[i] <- RMSE}
                                   return(results)}
```

# Determine best [dd] function construction
```{r dd}
selectBestDD1 <- function(TRAINX, TESTX, TRAINCOORDS, TESTCOORDS) {dd_values <- c(2, 3, 5)
                                                                   results <- CompareSP1(TRAINX, TESTX, TRAINCOORDS, TESTCOORDS, dd_values)
                                                                   best_index <- which.min(results$MAE + results$RMSE)
                                                                   best_dd <- results$dd[best_index]
                                                                   return(best_dd)}
```

# Determine best [dd] per covariate per country
```{r}
BestDD <- list()

for (df in names(Kriggdata)) {for (Variable in c("EFI", "RDI")) {
    
                                   if (Variable %in% names(Kriggdata[[df]]$Trainset)) {trainsets <- Kriggdata[[df]]$Trainset[[Variable]]
                                                                                       testsets <- Kriggdata[[df]]$Testset[[Variable]]
                                                                                       trainCC <- Kriggdata[[df]]$TrainCC[[Variable]]
                                                                                       testCC <- Kriggdata[[df]]$TestCC[[Variable]]
                                                                                        
                                                                                         train_data <- data.frame(trainsets, trainCC, 
                                                                                                                  stringsAsFactors = FALSE)
                                                                                         test_data <- data.frame(testsets, testCC,
                                                                                                                 stringsAsFactors = FALSE)
                                                                                         
                                                                                         train_data <- na.omit(train_data)
                                                                                         test_data <- na.omit(test_data)
                                                                                         
                                                                                         trainsets <- train_data[, 1] 
                                                                                         trainCC <- train_data[, 2:3]   
                                                                                         testsets <- test_data[, 1]   
                                                                                         testCC <- test_data[, 2:3] 
                                                                                        
                                  if (any(is.na(trainsets)) || any(is.na(testsets)) || any(is.na(trainCC)) || any(is.na(testCC))) {
                                      warning(paste("Missing values detected in", Variable, "for", df))
                                      next }
                                                                                        
                                       BestDD[[paste(df, Variable, sep = "_")]] <- selectBestDD1(trainsets,
                                                                                                 testsets,
                                                                                                 trainCC,
                                                                                                 testCC)}}}

rm(testsets, trainsets, Variable, df, testCC, trainCC, train_data, test_data, CompareSP1, selectBestDD1)
```

# Base raster preparation
```{r base-raster}
ADM0 <- st_read("Input/ADM0.gpkg") %>% 
        dplyr::select(admin0, name, geom) %>%
        rename("geometry" = geom) %>%
        dplyr::filter(!(toupper(admin0) %in% c("KY", "LB", "NI"))) %>%
        dplyr::mutate(admin0 = ifelse(tolower(name) == "moldova", "MB",
                               ifelse(tolower(name) == "dominican republic", "DR",
                               ifelse(tolower(name) == "india", "IA",
                               ifelse(tolower(name) == "liberia", "LB",
                               ifelse(tolower(name) == "madagascar", "MD",
                               ifelse(tolower(name) == "kyrgyzstan", "KY",
                               ifelse(tolower(name) == "namibia", "NM",
                               ifelse(tolower(name) == "burundi", "BU",
                               ifelse(tolower(name) == "guatemala", "GU",
                               ifelse(tolower(name) == "niger", "NI", toupper(admin0)))))))))))) %>%
        dplyr::select(admin0, geometry)

for (df in Dataframes) {DF <- ADM0 %>% dplyr::filter(admin0 == df)
                        write_stars(st_rasterize(st_as_sf(DF$geometry[1])), 
                                    paste0("Output/",
                                           df, ".tif"))
                        Raster <- raster(paste0("Output/",
                                             df, ".tif"))
                        Raster <- projectRaster(Raster, crs = crs(Raster), res = c(5/111, 5/85), ext = extent(Raster))
                        
                        writeRaster(Raster, paste0("Output/",
                                                   df, ".tif"), overwrite = TRUE) 
                        
                        assign(paste0(df, "R"), 
                               raster(paste0("Output/",
                                             df, ".tif")), envir = .GlobalEnv)}

rm(ADM0, df, DF, Raster)
```

# Krigging function construction for rasters
```{r krig-raster}
# For proportions [Must be 0-1 bound]                   # Raster
KrigingTIF1 <- function(X, COORDS, TIFNAME, dd, Raster){aa <- Raster
                                                        bb <- coordinates(aa)
                                                        ID <- is.na(getValues(aa))
                                                        miss <- which(ID == TRUE)
                                                        nonmiss <- which(ID == FALSE)
                                                
                                                        # Adjustment to extreme values to prevent errors
                                                        for(i in 1:length(X)){if(X[i] == 0){X[i] = 0 + 0.0001}	
                                                                              if(X[i] == 1){X[i] = 1 - 0.0001}}
                                                
                                                        # Logit transformation
                                                        logittrans <- log(X/(1-X))
                                                
                                                        dists	<- as.numeric(summary(dist(COORDS))[dd])
                                                        krig.fit <- Krig(COORDS, logittrans, theta = dists)
                                                                                                
                                                        # Subset dataset [too large otherwise]
                                                        cutsize <- 10000
                                                        rows <- dim(bb)[[1]]
                                                        cuts <- split(bb[, 1:2],
                                                                      rep(1:ceiling(rows / cutsize),    
                                                                          each = cutsize,
                                                                          length.out = rows))
                                                
                                                        # Predict
                                                        pred.out <- NA
                                                        pred.krig <- list()
                                                        backtrans	<- list()
                                                
                                                        for(ff in 1:length(cuts)){pred.krig[[ff]] <- as.vector(predict(krig.fit,matrix(cuts[[ff]], ncol = 2)))
                                                                                  backtrans[[ff]] <- as.vector(exp(pred.krig[[ff]]) / (1 + exp(pred.krig[[ff]])))}
                                                
                                                        # Combine subset predictions
                                                        pred.out <- unlist(backtrans)
                                                        pred.out[miss] <- NA
                                                        values(aa) <- pred.out
                                                
                                                        # Vizualize & export raster
                                                        plot(aa)
                                                        writeRaster(aa, TIFNAME, overwrite = TRUE)}
```

# Generate interpolated rasters
```{r interpolate}
for (df in Dataframes) {DF <- get(df)
                        for (Variable in c("EFI", "RDI")) {
                        Column <- paste0(Variable, "_prop")
                        if (Column %in% names(DF)) {DF <- DF[!is.na(DF[[Column]]), ]
                                                    Raster <- get(paste0(df, "R"), envir = .GlobalEnv)
                                                    DD <- BestDD[[paste0(df, "_", Variable)]]
                                                    
                                                    KrigingTIF1(DF[[Column]], 
                                                                DF[(ncol(DF) - 1):ncol(DF)], 
                                                                paste0("Output/",
                                                                       df, "_", Variable, ".tif"),
                                                                DD,
                                                                Raster)}}}

rm(KrigingTIF1, df, DF, Variable, Raster, DD, BestDD, Kriggdata)
```

# Extract coordinates & values per country-specific raster
```{r coords}
Rasters <- list()

for (df in Dataframes) {DF <- get(df)
                        for (Variable in c("EFI", "RDI")) {
                             Column <- paste0(Variable, "_prop")
                             if (Column %in% names(DF)) {
                                 Raster <- raster(paste0("Output/",
                                                         df, "_", Variable, ".tif"))
                                 Raster <- as.data.frame(Raster, xy = TRUE)
                                 names(Raster)[3] <- "Value" 
                                 Rasters[[paste0(Variable, "_",df)]] <- Raster}}}
```

# Merge data per covariate
```{r merge3}
GlobalData <- function(Variable) {bind_rows(lapply(Filter(function(x) grepl(Variable, x), names(Rasters)), 
                                                   function(df) Rasters[[df]]))}

EFI <- GlobalData("EFI") %>% dplyr::filter(!is.na(Value))
RDI <- GlobalData("RDI") %>% dplyr::filter(!is.na(Value))

rm(df, DF, Raster, Column, GlobalData, Rasters)
```

# Generate & export global rasters 
```{r global}
for (Variable in c("EFI", "RDI")) {
     DF <- get(Variable)
     if (all(c("x", "y") %in% colnames(DF))) {r <- raster()
                                              extent(r) <- extent(DF[, c("x", "y")])
                                              res(r) <- c(5/111, 5/85)
                                              Raster <- rasterize(DF[, c("x", "y")], r, field = DF$Value, fun = mean)
                                              assign(paste0(Variable, "R"), Raster)
                                                                     
                                              writeRaster(Raster, 
                                                          filename = paste0("Output/",
                                                                            Variable, "R.tif"),
                                                          format = "GTiff",
                                                          overwrite = TRUE)
                                              
                                              Palette <- colorNumeric(palette = c("#799AE3", "#0F1D41"), 
                                                                      domain = c(min(values(Raster), na.rm = TRUE),
                                                                                 max(values(Raster), na.rm = TRUE) + 0.001),
                                                                      na.color = "transparent")
                                                                     
                                              Map <- leaflet() %>%
                                                     addProviderTiles("CartoDB.Positron") %>%
                                                     addRasterImage(Raster, opacity = 0.8, 
                                                                    colors = Palette) %>%
                                                     addLegend("bottomright", pal = Palette, opacity = 0.8, title = paste0(Variable),
                                                               values = c(0, 1))
                                              
                                             print(Map)} 
     
     else {warning(paste("Data frame for", Variable, "does not have the required coordinate columns and is skipped."))}}

rm(list = setdiff(ls(), c("EFIR", "RDIR")))
```

# Population
```{r population}
Population <- raster("Input/global_pop_2026_CN_1km_R2025A_UA_v1.tif")

Raster <- raster(extent(Population), res = c(5/111, 5/85), crs = crs(Population))

Population <- resample(Population, Raster, method = "bilinear")
Population <- crop(Population, Raster)
```

# Malaria
```{r malaria}
Malaria <- raster("Input/202206_Global_Pf_Incidence_Rate_2020.tif")
Malaria <- resample(Malaria, Raster, method = "bilinear")
Malaria <- crop(Malaria, Raster)
```

# Urbanization
```{r urbanization, warning = FALSE}
Urbanization <- raster(do.call(merge, lapply(files <- list.files("Input/Urbanization",
                                                                 pattern = "\\.(tif)$", full.names = TRUE, ignore.case = TRUE), rast)))

Urbanization <- projectRaster(from = Urbanization, to = Raster, method = "ngb")
Urbanization <- crop(Urbanization, Raster)
```

# Consanguinity
```{r consanguinity}
ADM0 <- st_read("Input/ADM0.gpkg") %>% 
        mutate(ISO = countrycode(name, origin = "country.name", destination = "iso3c"),
               ISO = ifelse(is.na(ISO) & name == "Kosovo", "XKX", ISO)) %>%
        dplyr::select(ISO, geom)

Consanguinity <- read_excel("Input/Global Consanguinity.xlsx") %>%
                 mutate(ISO = countrycode(Country, origin = "country.name", destination = "iso3c")) %>%
                 dplyr::select(ISO, Consanguinity)

Consanguinity <- merge(Consanguinity, ADM0, by = "ISO") %>%
                 dplyr::select(-ISO) %>%
                 st_as_sf()

Consanguinity <- rasterize(x = Consanguinity,
                           y = raster(extent(Raster), res = res(Raster), crs = crs(Raster)), 
                           field = as.numeric(Consanguinity$Consanguinity), 
                           fun = mean, 
                           background = NA)
```

# EFI
```{r efi}
EFI <- read_excel("Input/Global EFI (HIEF).xlsx") %>%
       group_by(Country) %>%
       dplyr::filter(Year == max(Year)) %>%
       ungroup() %>%
       mutate(Country = ifelse(Country == "Yemen Arab Republic", "Yemen", Country),
              ISO = countrycode(Country, origin = "country.name", destination = "iso3c")) %>%
       dplyr::filter(!is.na(ISO)) %>%
       dplyr::filter(Year == 2013) %>%
       dplyr::select(ISO, EFI)

EFI <- merge(EFI, ADM0, by = "ISO") %>%
       dplyr::select(-ISO) %>%
       st_as_sf()

EFI <- rasterize(x = EFI,
                 y = raster(extent(Raster), res = res(Raster), crs = crs(Raster)), 
                 field = as.numeric(EFI$EFI), 
                 fun = mean, 
                 background = NA)
```

# RDI
```{r rdi}
RDI <- read_excel("Input/Global Religious Diversity Index (RDI).xlsx") %>%
       dplyr::select(Country, RDI) %>%
       mutate(ISO = countrycode(Country, origin = "country.name", destination = "iso3c"),
              ISO = ifelse(is.na(ISO) & Country == "Kosovo", "XKX",
                    ifelse(is.na(ISO) & Country == "Netherlands Antilles", "ANT",
                    ifelse(is.na(ISO) & Country == "Channel Islands", "CHANNEL", ISO)))) %>%
       dplyr::select(ISO, RDI)

RDI <- merge(RDI, ADM0, by = "ISO") %>%
       dplyr::select(-ISO) %>%
       st_as_sf()

RDI <- rasterize(x = RDI,
                 y = raster(extent(Raster), res = res(Raster), crs = crs(Raster)), 
                 field = as.numeric(RDI$RDI), 
                 fun = mean, 
                 background = NA)
```

# Update rasters based on interpolated data
```{r update-raster}
EFI <- overlay(EFI, resample(EFIR, EFI, method = "bilinear"), fun = function(x, y) {ifelse(!is.na(y), y, x)})
RDI <- overlay(RDI, resample(RDIR, RDI, method = "bilinear"), fun = function(x, y) {ifelse(!is.na(y), y, x)})
```

# Covariates
```{r assemble1}
Covariates <- stack(Population, Malaria, Urbanization, Consanguinity, EFI, RDI)
names(Covariates) <- c("Population", "Malaria", "Urbanization", "Consanguinity", "EFI", "RDI")

rm(list = setdiff(ls(), c("Covariates", "Raster")))
```

# Counselling
```{r counselling}
ADM0 <- st_read("Input/ADM0.gpkg") %>% 
        mutate(ISO = countrycode(name, origin = "country.name", destination = "iso3c"),
               ISO = ifelse(is.na(ISO) & name == "Kosovo", "XKX", ISO)) %>%
        dplyr::select(ISO, geom)

Counselling <- read_excel("Input/IthaMaps service availability.xlsx") %>%
               mutate(Country = ifelse(Country == "Leba0n", "Lebanon",
                                ifelse(Country == "0rway", "Norway", Country)),
                      ISO = countrycode(Country, origin = "country.name", destination = "iso3c")) %>%
               dplyr::filter(Service == "Genetic counselling") %>%
               dplyr::select(ISO, Availability) %>%
               rename(Counselling = Availability)

Counselling <- merge(Counselling, ADM0, by = "ISO") %>%
               dplyr::select(-ISO) %>%
               st_as_sf()

Counselling <- rasterize(x = Counselling,
                         y = raster(extent(Raster), res = res(Raster), crs = crs(Raster)), 
                         field = as.numeric(Counselling$Counselling), 
                         fun = mean, 
                         background = NA)

Counselling2 <- resample(Counselling, Raster, method = "bilinear")
Counselling2 <- crop(Counselling2, Raster)

Counselling <- overlay(Counselling, Counselling2, fun = function(x, y) {ifelse(!is.na(y), y, x)})
```

# Covariates
```{r assemble2}
Covariates <- stack(Covariates, Counselling)
names(Covariates) <- c("Population", "Malaria", "Urbanization", "Consanguinity", "EFI", "RDI", "Counselling")

rm(list = setdiff(ls(), "Covariates"))
```

# Prevention
```{r prevention}
ADM0 <- st_read("Input/ADM0.gpkg") %>% 
        mutate(ISO = countrycode(name, origin = "country.name", destination = "iso3c"),
               ISO = ifelse(is.na(ISO) & name == "Kosovo", "XKX", ISO)) %>%
        dplyr::select(ISO, geom)

Prevention <- read_excel("Input/IthaMaps service availability.xlsx") %>%
              mutate(Country = ifelse(Country == "Leba0n", "Lebanon",
                               ifelse(Country == "0rway", "Norway", Country)),
                     ISO = countrycode(Country, origin = "country.name", destination = "iso3c")) %>%
              dplyr::filter(Service == "Prevention program") %>%
              dplyr::select(ISO, Availability) %>%
              rename(Prevention = Availability)

Prevention <- merge(Prevention, ADM0, by = "ISO") %>%
              dplyr::select(-ISO) %>%
              st_as_sf()

Prevention <- rasterize(x = Prevention,
                        y = raster(extent(Raster), res = res(Raster), crs = crs(Raster)), 
                        field = as.numeric(Prevention$Prevention), 
                        fun = mean, 
                        background = NA)

Prevention2 <- resample(Prevention, Raster, method = "bilinear")
Prevention2 <- crop(Prevention2, Raster)

Prevention <- overlay(Prevention, Prevention2, fun = function(x, y) {ifelse(!is.na(y), y, x)})
```

# Covariates
```{r assemble3}
Covariates <- stack(Covariates, Prevention)
names(Covariates) <- c("Population", "Malaria", "Urbanization", "Consanguinity", "EFI", "RDI", "Counselling", "Prevention")

writeRaster(Covariates, filename = "Output/Covariates3.tiff", 
            format = "GTiff", 
            overwrite = TRUE)

rm(list = setdiff(ls(), "Covariates"))
```